---
title: "bvarnet"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{bvarnet}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This package implements methods to analyse (multilevel) Vector Autoregression commonly used for the analysis of intensive longitudinal datasets like EMA data.

In the following vignette we will introduce the basic functionality of the package. For more elaborated usecases, we refer to the additional vignettes

- `vignette("Mixed-Model")`
- `vignette("Missing-Data")`
- `vignette("Random-Effect")`
- `vignette("Prediction")`

First, we have to load the package:

```{r setup}
library(bvarnet)
```

There is some missing data in the dataset. The models default options handle this by themselves. For a further elaboration on this, you can read `vignette("Missing-Data")`.

```{r}
data <- openesm::get_dataset("0004")

df <- data$data

df$anxious <- floor(df$anxious)
df$calm <- floor(df$calm)
df$conventional <- floor(df$conventional)
df$critical <- floor(df$critical)
df$dependable <- floor(df$dependable)
```


```{r}
fit <- bvar(
  id_col = "id",
  time_col = "day",
  y_cols = c("anxious", "calm", "conventional", "critical", "dependable"),
  x_cols = NULL,
  re_cols = NULL,
  re_temporal = FALSE,
  K = 1,
  data = df,
  family = c("ordinal"),
  priors = set_priors(),
  iter = 4000,
  warmup = 1000,
  chains = 4,
  cores = 4,
  seed = 1337
)
```

```{r}
summary(fit)
print(fit)
fit

params <- extract_param(fit)

extract_temporal(fit)
extract_network_matrix(fit)
```

```{r}
summary(fit)
```

