---
title: "Hypothesis-Testing"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Hypothesis-Testing}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

has_cmdstan <- instantiate::stan_cmdstan_exists()
```

First, we have to load the package:

```{r setup}
library(bvarnet)

# subject to be removed again...
library(openesm)
library(qgraph)
```

There is some missing data in the dataset. The models default options handle this by themselves. For a further elaboration on this, you can read `vignette("Missing-Data")`.

```{r}
data <- get_dataset("0004")

df <- data$data

df$anxious <- floor(df$anxious)
df$calm <- floor(df$calm)
df$conventional <- floor(df$conventional)
df$critical <- floor(df$critical)
df$dependable <- floor(df$dependable)

library(dplyr)
df_mis <- na.omit(df %>% select(id, day, anxious, calm, conventional, critical, dependable, sleep_hour))
```

This time, well simply use the default priors, for the ordinal model they are

```{r, eval = has_cmdstan}
fit <- bvar(
  id_col = "id",
  time_col = "day",
  y_cols = c("anxious", "calm", "conventional", "critical", "dependable"),
  x_cols = c("sleep_hour"),
  re_cols = NULL,
  re_temporal = FALSE,
  K = 1,
  data = df,
  family = c("ordinal"),
  priors = set_priors(),
  iter = 4000,
  warmup = 1000,
  chains = 4,
  cores = 4,
  seed = 1337
)
```


Again, lets get the model overview and summary:
```{r, eval = has_cmdstan}
print(fit)
summary(fit)
```

To estimate Bayes factors for a single parameter, or a set of parameters, we can use the `bf_table()` function.

To perform a hypothesis test on effects involving a specific variable, we can call the `bf_table()` function, specifying which variable we want to look at using the `variable` argument. Lets investigate if the hours of sleep the individual got in the previous night (`sleep_hour`) has an influence on the network structure:

```{r}
bf_tab <- bf_table(fit, variable = "sleep_hour")
bf_tab
```

The output now shows 6 rows. The five "Fixed Effect" rows show if the the covariate `sleep_hour` has an influence on each node separately. The "Fixed Effect (join)" row displays a single Bayes factor that in this case tests the hypothesis

H: Is there a difference in the baseline of the network structure that is explained by the variable `sleep_hour`

As we can see, the `BF_{10} = bf_tab[]`. Therefore we can conlude that the variable `sleep_hour` does not influence the baseline level of the network.


For more elaborate analyses, or to report more results at once, we can get a table with all single and joint Bayes factors. For this we can call the `bf_table(fit)` function

```{r}
bf_tab <- bf_table(fit)
```

The table can be used to easily get a set of Bayes factors. If we have hypotheses concerning the temporal structure, we can call the `bf_table(fit)` function using the argument `type = "temporal"`. The following arguments are available for the `type` argument: ""



$\mathbf{x}_{it}^ = \mathbf{M}^T \mathbf{x}_{it}$
