---
title: "Missing-Data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Missing-Data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

First, we have to load the package:

```{r setup}
library(bvarnet)

# subject to be removed again...
library(openesm)
library(qgraph)
```

There is some missing data in the dataset. The models default options handle this by themselves. For a further elaboration on this, you can read `vignette("Missing-Data")`.

```{r}
data <- get_dataset("0004")

df <- data$data

df$anxious <- floor(df$anxious)
df$calm <- floor(df$calm)
df$conventional <- floor(df$conventional)
df$critical <- floor(df$critical)
df$dependable <- floor(df$dependable)
```

Inspecting the dataset, we can see that there are quite a lot of missing values:
```{r}

```



```{r, eval = has_cmdstan}
fit <- bvar(
  id_col = "id",
  time_col = "day",
  y_cols = c("anxious", "calm", "conventional", "critical", "dependable"),
  x_cols = NULL,
  re_cols = NULL,
  re_temporal = FALSE,
  K = 1,
  data = df,
  family = c("ordinal"),
  priors = set_priors(),
  iter = 4000,
  warmup = 1000,
  chains = 4,
  cores = 4,
  seed = 1337
)
```

We can print an overview of the model we just ran using the `print(fit)` functions:
```{r, eval = has_cmdstan}
print(fit)
```

Here we get information about the number of variables and lags, the number of observations and some first indications if the model did converge.

To further inspect the model parameters we can use the `summary(fit)` function:

```{r, eval = has_cmdstan}
summary(fit)
```
