---
title: "Mixed-Model"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Mixed-Model}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

has_cmdstan <- instantiate::stan_cmdstan_exists()
```

```{r setup}
library(bvarnet)
library(qgraph)
```

## Mixed-family VAR

`bvarnet` allows combining different outcome families in a single VAR model
by passing a named vector to the `family` argument. This is useful when the
variables in the network are measured on different scales.

In this example we fit a mixed-family VAR(1) with three types of outcome:

- **ordinal**: `anxious` and `calm` (5-point Likert items)
- **binary**: `happyornot` (binarised daily average)
- **gaussian**: `sleep_duration` (hours)

## Load Data
Again, we start with loading the data
```{r}
data(studentlife)
```


### Priors

For a mixed-family model we specify priors for all parameter types that can
appear. Here that means `phi` (temporal), `beta` (intercept / fixed effects),
`kappa` (ordinal thresholds), and `sigma` (gaussian residual SD).

```{r}
priors <- set_priors(
  phi   = prior(family = "normal", loc = 0, scale = 0.1),
  beta  = prior(family = "normal", loc = 0, scale = 1),
  kappa = prior(family = "normal", loc = 0, scale = 1),
  sigma = prior(family = "normal", loc = 0, scale = 1)
)
```

### Estimation

We pass a named character vector to `family` so that each outcome gets the
appropriate likelihood.

```{r, eval = has_cmdstan}
#potential bug: we were able to run with binary that were saved as numeric (and have decimals)
fit_mixed <- bvar(
  id_col   = "id",
  time_col = "day",
  y_cols   = c("stress_level", "happyornot", "sleep_duration"),
  x_cols   = NULL,
  re_cols  = NULL,
  re_temporal = FALSE,
  K        = 1,
  data     = studentlife,
  family   = c(stress_level = "ordinal",
               happyornot = "bernoulli",
               sleep_duration = "gaussian"),
  priors   = priors,
  iter     = 4000,
  warmup   = 2000,
  chains   = 4,
  cores    = 4,
  seed     = 1337
)
```

### Results

```{r, eval = has_cmdstan}
print(fit_mixed)
```

```{r, eval = has_cmdstan}
summary(fit_mixed)
```

The temporal network can be extracted and inspected as usual:

```{r, eval = has_cmdstan}
nw_mat <- extract_network_matrix(fit_mixed)
qgraph(nw_mat)
```
