---
title: "Prediction"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Prediction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


```{r setup}
library(bvarnet)

# subject to be removed again...
library(qgraph)
```

## Load Data

```{r}
data(studentlife)
```

## Fit the model

To show the full capabilities of the prediction module, we will use the multilevel VAR model as example:

```{r, eval = has_cmdstan}
fit <- bvar(
  id_col = "id",
  time_col = "day",
  y_cols = c("anxious", "calm", "conventional", "critical", "dependable"),
  x_cols = NULL,
  re_cols = c("Intercept"),
  re_temporal = TRUE,
  K = 1,
  data = studentlife,
  family = c("ordinal"),
  iter = 4000,
  warmup = 1000,
  chains = 4,
  cores = 4,
  seed = 1337
)
```

```{r}
summary(fit)
```

## 2 — In-Sample Fitted Values

The simplest call returns one-step-ahead fitted values on the linear predictor scale for the training data:

```{r}
p <- predict(fit)
p
```




This returns a **matrix** with dimensions `[nrow(original_data), p]`, where `p` is the number of outcome nodes. The first `K` rows per subject are `NA` because those time points lack sufficient lag history.

