---
title: "Prediction and Model Calibration"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Prediction and Model Calibration}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse  = TRUE,
  comment   = "#>",
  fig.width = 7,
  fig.height = 5
)
has_cmdstan <- instantiate::stan_cmdstan_exists()
```

## Overview

The `predict()` method for **bvarnet** objects generates fitted values for all
three response families. The `type` argument controls the scale:

| Family | `type = "link"` | `type = "response"` | `type = "probabilities"` |
|--------|----------------|---------------------|--------------------------|
| `"ordinal"` | latent linear predictor | modal category | category probabilities (list of data frames) |
| `"gaussian"` | fitted mean (= response scale) | fitted mean | — |
| `"bernoulli"` | log-odds | predicted probability (0–1) | — |

Throughout this vignette we use two freely available ESM datasets from the
**openESM** database (Siepe 2025) accessed via the `openesm` package.
The `openesm::get_dataset()` calls will serve data from a local cache when the
dataset has been downloaded before, and download it from Zenodo on first use.

**Dataset selection criterion.** A multilevel VAR prediction plot is only
informative if the outcomes carry genuine *within-person* temporal signal —
that is, if knowing yesterday's value (after removing each person's mean) helps
predict today's. We selected variables with within-person lag-1 autocorrelation
$r \geq 0.25$ on the ordinal/binary scale, and $r \geq 0.55$ on the Gaussian
scale.

---

## 1 — Ordinal VAR: COVID Lockdown Affect (Fried 2021)

We use the **Fried (2021)** dataset: 79 university students assessed 4 times/day
over 14 days during the first COVID lockdown. Emotional states were rated on a
5-point scale (1 = *Not at all*, 5 = *Extremely*).

We model four items with the highest within-person lag-1 autocorrelation
(range: 0.27 – 0.32): `alone`, `covid_worry`, `nothing_look_forward`, and
`anhedonia`. These tap distinct but interrelated facets of the COVID pandemic
experience — social isolation, pandemic-related worry, anticipatory anhedonia,
and inability to enjoy activities.

```{r load-ordinal, eval = has_cmdstan, message = FALSE, warning = FALSE}
library(bvarnet)
library(openesm)

# Download / load from cache
fried_raw <- get_dataset("0001_fried")$data

y_vars <- c("alone", "covid_worry", "nothing_look_forward", "anhedonia")

fried <- fried_raw[complete.cases(fried_raw[, y_vars]), ]
fried$id  <- as.character(fried$id)
fried$day <- as.integer(fried$day)

cat("Dimensions:", nrow(fried), "rows ×", ncol(fried), "cols\n")
cat("Subjects:", length(unique(fried$id)), "| Obs. per item range:",
    min(colSums(!is.na(fried[, y_vars]))), "–",
    max(colSums(!is.na(fried[, y_vars]))), "\n")
cat("Response scale: 1 – 5 (Not at all … Extremely)\n")
cat("Within-person lag-1 r (approx.):",
    "alone ≈ 0.32, covid_worry ≈ 0.29,",
    "nothing_look_forward ≈ 0.27, anhedonia ≈ 0.27\n")
```

### Fit the model

```{r fit-ordinal, eval = has_cmdstan, message = FALSE, warning = FALSE}
fit_ord <- bvar(
  id_col      = "id",
  time_col    = "day",
  y_cols      = y_vars,
  x_cols      = NULL,
  re_cols     = c("Intercept"),
  re_temporal = FALSE,
  K           = 1,
  data        = fried,
  family      = "ordinal",
  iter        = 4000,
  warmup      = 1000,
  chains      = 4,
  cores       = 4,
  seed        = 42,
  save_data   = TRUE
)
```

```{r summary-ordinal, eval = has_cmdstan}
summary(fit_ord)
```

### 1a — What does `predict()` return for ordinal models?

The default call returns one-step-ahead fitted values on the **latent linear
predictor** scale. For ordinal models these are not directly interpretable, so
we move straight to category probabilities.

```{r predict-probs-ordinal, eval = has_cmdstan}
probs_ord <- predict(fit_ord, type = "probabilities")
# probs_ord is a named list: one data.frame per outcome
# Each data.frame has columns id, day, cat_1 … cat_5
head(probs_ord[["alone"]])

# Shared constants used in all ordinal plots
n_cats   <- 5L
cat_cols <- paste0("cat_", seq_len(n_cats))
ord_pal  <- hcl.colors(n_cats, "Blues 3")
```

### 1b — Calibration: Are the right categories predicted?

A well-calibrated model assigns the **highest probability to the observed
category**. The barplot shows, for each observed rating (x-axis), the mean
predicted probability per category (stacked bars). A diagonal pattern — low
bars dominated by dark blue at the left, shifting to light blue at the right —
indicates the model tracks ordinal responses.

```{r calibration-barplot, eval = has_cmdstan, fig.height = 7}
obs_ord <- fit_ord$data_used

par(mfrow = c(2, 2), mar = c(4, 4, 2.5, 1))
for (y in y_vars) {
  cmp <- merge(
    obs_ord[, c("id", "day", y)],
    probs_ord[[y]],
    by = c("id", "day")
  )
  cal <- matrix(NA_real_, n_cats, n_cats,
                dimnames = list(observed  = seq_len(n_cats),
                                predicted = cat_cols))
  for (k in seq_len(n_cats)) {
    rows_k <- cmp[[y]] == k
    if (any(rows_k)) cal[k, ] <- colMeans(cmp[rows_k, cat_cols, drop = FALSE])
  }
  barplot(
    t(cal), beside = FALSE, col = ord_pal,
    xlab = "Observed category", ylab = "Mean predicted probability",
    main = y, ylim = c(0, 1)
  )
  legend("topright",
         legend = rev(seq_len(n_cats)), title = "Predicted cat.",
         fill = rev(ord_pal), cex = 0.7, bty = "n")
}
```

### 1c — Individual probability trajectory

The stacked area chart shows how predicted category probabilities evolve across
days for the participant with the highest day-to-day variability in `alone`.
Black dots show the rescaled observed rating (1 → 0, 5 → 1); a dot inside its
corresponding coloured band indicates a correct modal prediction.

```{r trajectory, eval = has_cmdstan, fig.height = 5}
sd_per_subj <- tapply(obs_ord$alone, obs_ord$id, sd, na.rm = TRUE)
subj_pick   <- names(which.max(sd_per_subj))

y <- "alone"
cmp <- merge(
  obs_ord[obs_ord$id == subj_pick, c("id", "day", y)],
  probs_ord[[y]][probs_ord[[y]]$id == subj_pick, ],
  by = c("id", "day")
)
cmp <- cmp[order(cmp$day), ]

plot(NULL, xlim = range(cmp$day), ylim = c(0, 1),
     xlab = "Day", ylab = "Predicted probability",
     main = paste0(y, " — participant ", subj_pick))

cum <- numeric(nrow(cmp))
for (k in seq_len(n_cats)) {
  new_cum <- cum + cmp[[paste0("cat_", k)]]
  polygon(c(cmp$day, rev(cmp$day)), c(new_cum, rev(cum)),
          col = ord_pal[k], border = NA)
  cum <- new_cum
}
points(cmp$day, (cmp[[y]] - 1) / (n_cats - 1), pch = 16, cex = 1.2)
legend("topright", legend = rev(seq_len(n_cats)), title = "Category",
       fill = rev(ord_pal), cex = 0.7, bty = "n")
```

### 1d — Observed vs predicted category frequencies over time

We compare observed category proportions (solid lines) with mean predicted
probabilities (dashed) aggregated across all subjects at each time point.
Close tracking indicates the model captures population-level temporal dynamics.

```{r freq-over-time, eval = has_cmdstan, fig.height = 7}
par(mfrow = c(2, 2), mar = c(4, 4, 2.5, 1))
for (y in y_vars) {
  cmp  <- merge(obs_ord[, c("id", "day", y)], probs_ord[[y]], by = c("id", "day"))
  days <- sort(unique(cmp$day))
  obs_m  <- matrix(NA_real_, length(days), n_cats)
  pred_m <- matrix(NA_real_, length(days), n_cats)
  for (i in seq_along(days)) {
    rows_d   <- cmp$day == days[i]
    obs_cats <- cmp[[y]][rows_d]
    for (k in seq_len(n_cats)) {
      obs_m[i, k]  <- mean(obs_cats == k, na.rm = TRUE)
      pred_m[i, k] <- mean(cmp[[paste0("cat_", k)]][rows_d], na.rm = TRUE)
    }
  }
  ymax <- max(obs_m, pred_m, na.rm = TRUE)
  plot(NULL, xlim = range(days), ylim = c(0, ymax),
       xlab = "Day", ylab = "Proportion", main = y)
  for (k in seq_len(n_cats)) {
    lines(days, obs_m[, k],  col = ord_pal[k], lwd = 2, lty = 1)
    lines(days, pred_m[, k], col = ord_pal[k], lwd = 2, lty = 2)
  }
  legend("topright",
         legend = c(paste0("Cat. ", seq_len(n_cats)), "Observed", "Predicted"),
         col    = c(ord_pal, "black", "black"),
         lty    = c(rep(NA, n_cats), 1, 2),
         pch    = c(rep(15, n_cats), NA, NA),
         lwd = 2, cex = 0.55, bty = "n", ncol = 2)
}
```

### 1e — Modal category accuracy

The proportion of observations for which the predicted modal category matches
the observed response. The majority-class baseline reflects always predicting
the most common category — a stronger reference than uniform chance when the
distribution is skewed.

```{r modal-accuracy, eval = has_cmdstan}
acc_tbl <- data.frame(
  variable     = y_vars,
  pct_correct  = NA_real_,
  pct_majority = NA_real_
)
for (i in seq_along(y_vars)) {
  y   <- y_vars[i]
  cmp <- merge(obs_ord[, c("id", "day", y)], probs_ord[[y]], by = c("id", "day"))
  modal_pred              <- max.col(cmp[, cat_cols])
  acc_tbl$pct_correct[i]  <- round(100 * mean(modal_pred == cmp[[y]], na.rm = TRUE), 1)
  majority_cat            <- as.integer(names(which.max(table(cmp[[y]]))))
  acc_tbl$pct_majority[i] <- round(100 * mean(majority_cat == cmp[[y]], na.rm = TRUE), 1)
}
names(acc_tbl) <- c("Variable", "% correct (model)", "% correct (majority baseline)")
knitr::kable(acc_tbl, align = "lrr")
```

---

## 2 — Gaussian VAR: Daily Behavioral Rhythms from Phone Sensing (Wang 2020)

We use the **Wang (2020)** dataset: 49 university students passively monitored
over approximately six weeks via their smartphones. Each day yields two
continuous behavioral measures derived from passive phone sensing:

| Variable | Description | Lag-1 *r* |
|---|---|---|
| `loc_visit_num_ep_0` | Number of distinct locations visited (full day, from GPS) | 0.68 |
| `audio_convo_num_ep_1` | Number of conversational audio episodes (morning period) | 0.58 |

Both variables have strong within-person autocorrelation because behavioral
routines (commuting, socialising) are stable across consecutive days. The data
are one row per subject–day (no within-day beeps to deduplicate).

```{r load-gaussian, eval = has_cmdstan, message = FALSE, warning = FALSE}
wang_raw <- get_dataset("0004_wang")$data

y_gauss <- c("loc_visit_num_ep_0", "audio_convo_num_ep_1")

wang <- wang_raw[complete.cases(wang_raw[, y_gauss]), ]
wang$id  <- as.character(wang$id)
wang$day <- as.integer(wang$day)

cat("Dimensions:", nrow(wang), "rows ×", ncol(wang), "cols\n")
cat("Subjects:", length(unique(wang$id)), "| Days (median per subject):",
    median(as.integer(table(wang$id))), "\n")
cat("loc_visit_num_ep_0:   M =", round(mean(wang$loc_visit_num_ep_0), 2),
    " SD =", round(sd(wang$loc_visit_num_ep_0), 2), "\n")
cat("audio_convo_num_ep_1: M =", round(mean(wang$audio_convo_num_ep_1), 2),
    " SD =", round(sd(wang$audio_convo_num_ep_1), 2), "\n")
```

### Fit the model

```{r fit-gaussian, eval = has_cmdstan, message = FALSE, warning = FALSE}
fit_gauss <- bvar(
  id_col      = "id",
  time_col    = "day",
  y_cols      = y_gauss,
  x_cols      = NULL,
  re_cols     = c("Intercept"),
  re_temporal = FALSE,
  K           = 1,
  data        = wang,
  family      = "gaussian",
  iter        = 4000,
  warmup      = 1000,
  chains      = 4,
  cores       = 4,
  seed        = 42,
  save_data   = TRUE
)
```

```{r summary-gaussian, eval = has_cmdstan}
summary(fit_gauss)
```

### Observed vs predicted

For Gaussian outcomes the fitted values are on the response scale. Points close
to the identity line (red) indicate the model captures individual variation
beyond each person's mean level.

```{r gaussian-scatter, eval = has_cmdstan, fig.height = 5}
pred_gauss <- predict(fit_gauss, type = "response")
obs_gauss  <- fit_gauss$data_used
cmp_g      <- merge(obs_gauss, pred_gauss, by = c("id", "day"))

par(mfrow = c(1, 2), mar = c(4, 4, 2.5, 1))
for (y in y_gauss) {
  observed  <- cmp_g[[y]]
  predicted <- cmp_g[[paste0("predicted_", y)]]
  r2        <- cor(observed, predicted, use = "complete.obs")^2
  lab       <- sub("_ep_[0-9]", "", y)   # short label for axis
  plot(observed, predicted,
       xlab = paste("Observed", lab), ylab = paste("Predicted", lab),
       main = paste0(lab, "  (R\u00b2 = ", round(r2, 2), ")"),
       pch = 16, cex = 0.5, col = rgb(0.2, 0.4, 0.8, 0.35))
  abline(0, 1, col = "firebrick", lwd = 2)
}
```

The in-sample R² is an optimistic upper bound; out-of-sample evaluation would
give a more conservative estimate.

### Residual distribution

Histograms of residuals check whether the normality assumption is reasonable.

```{r gaussian-residuals, eval = has_cmdstan, fig.height = 4}
par(mfrow = c(1, 2), mar = c(4, 4, 2.5, 1))
for (y in y_gauss) {
  resid <- cmp_g[[y]] - cmp_g[[paste0("predicted_", y)]]
  lab   <- sub("_ep_[0-9]", "", y)
  hist(resid, breaks = 30, freq = FALSE,
       xlab = "Residual (observed \u2212 predicted)",
       main = paste0(lab, " residuals"),
       col = "steelblue3", border = "white")
  curve(dnorm(x, mean = 0, sd = sd(resid, na.rm = TRUE)),
        col = "firebrick", lwd = 2, add = TRUE)
}
```

---

## 3 — Binary VAR: Elevated Loneliness and Worry Days (Fried 2021)

We return to the **Fried (2021)** data and create binary indicators by
thresholding three of the ordinal items at the midpoint of the scale ("at or
above moderately"):

| Variable | Threshold | Meaning |
|---|---|---|
| `alone_elev` | `alone ≥ 3` | At or above "moderately" alone |
| `worry_elev` | `covid_worry ≥ 3` | At or above "moderately" worried |
| `anhedonia_elev` | `anhedonia ≥ 3` | At or above "moderately" unable to enjoy |

Thresholding at the scale midpoint identifies *clinically elevated* days. The
underlying ordinal items have within-person lag-1 r ≈ 0.27–0.32, so the binary
indicators retain meaningful temporal autocorrelation.

```{r prepare-binary, eval = has_cmdstan}
fried_bin              <- fried
fried_bin$alone_elev    <- as.integer(fried_bin$alone        >= 3)
fried_bin$worry_elev    <- as.integer(fried_bin$covid_worry  >= 3)
fried_bin$anhedonia_elev <- as.integer(fried_bin$anhedonia   >= 3)

y_bin <- c("alone_elev", "worry_elev", "anhedonia_elev")

cat("Base rates:\n")
for (y in y_bin) {
  cat(" ", y, ":", round(100 * mean(fried_bin[[y]], na.rm = TRUE), 1), "%\n")
}
```

### Fit the model

```{r fit-binary, eval = has_cmdstan, message = FALSE, warning = FALSE}
fit_bin <- bvar(
  id_col      = "id",
  time_col    = "day",
  y_cols      = y_bin,
  x_cols      = NULL,
  re_cols     = c("Intercept"),
  re_temporal = FALSE,
  K           = 1,
  data        = fried_bin,
  family      = "bernoulli",
  iter        = 4000,
  warmup      = 1000,
  chains      = 4,
  cores       = 4,
  seed        = 42,
  save_data   = TRUE
)
```

```{r summary-binary, eval = has_cmdstan}
summary(fit_bin)
```

### 3a — Reliability diagram (calibration)

A **reliability diagram** checks whether predicted probabilities are on the
right scale: observations in the 30–40% predicted bin should occur at a 30–40%
rate. All points on the 45° diagonal = perfect calibration.

```{r reliability-diagram, eval = has_cmdstan, fig.height = 5}
pred_bin <- predict(fit_bin, type = "response")
obs_bin  <- fit_bin$data_used
cmp_b    <- merge(obs_bin, pred_bin, by = c("id", "day"))

n_bins <- 10
par(mfrow = c(1, length(y_bin)), mar = c(4, 4, 3, 1))
for (y in y_bin) {
  observed  <- cmp_b[[y]]
  predicted <- cmp_b[[paste0("predicted_", y)]]

  breaks  <- seq(0, 1, length.out = n_bins + 1)
  bin_idx <- cut(predicted, breaks = breaks, include.lowest = TRUE, labels = FALSE)
  bin_mid <- tapply(predicted, bin_idx, mean, na.rm = TRUE)
  bin_obs <- tapply(observed,  bin_idx, mean, na.rm = TRUE)
  bin_n   <- tapply(observed,  bin_idx, length)
  cex_pts <- sqrt(bin_n / max(bin_n, na.rm = TRUE)) * 3 + 0.5

  plot(NULL, xlim = c(0, 1), ylim = c(0, 1),
       xlab = "Mean predicted probability",
       ylab = "Observed frequency",
       main = sub("_elev", "", y))
  abline(0, 1, col = "grey60", lty = 2)
  abline(h = mean(observed, na.rm = TRUE), col = "grey80", lty = 3)
  points(bin_mid, bin_obs, pch = 16, cex = cex_pts,
         col = rgb(0.2, 0.4, 0.8, 0.7))
  text(0.02, 0.97,
       paste0("Base rate: ", round(100 * mean(observed, na.rm = TRUE), 1), "%"),
       adj = c(0, 1), cex = 0.8, col = "grey30")
}
```

Point size is proportional to the number of observations in each bin. The
dashed diagonal is the calibration ideal; the dotted horizontal line shows the
marginal base rate.

### 3b — Discrimination: ROC curve

The ROC curve shows whether the model **ranks** positive days above negative
ones. AUC = 0.5 is chance; AUC = 1.0 is perfect discrimination.

```{r roc-curve, eval = has_cmdstan, fig.height = 5}
roc_points <- function(obs, pred) {
  thresholds <- sort(unique(pred), decreasing = TRUE)
  fpr <- tpr <- numeric(length(thresholds) + 2)
  fpr[1] <- tpr[1] <- 0
  for (i in seq_along(thresholds)) {
    pp  <- pred >= thresholds[i]
    tp  <- sum(pp  & obs == 1, na.rm = TRUE)
    fp  <- sum(pp  & obs == 0, na.rm = TRUE)
    fn  <- sum(!pp & obs == 1, na.rm = TRUE)
    tn  <- sum(!pp & obs == 0, na.rm = TRUE)
    tpr[i + 1] <- tp / (tp + fn)
    fpr[i + 1] <- fp / (fp + tn)
  }
  fpr[length(fpr)] <- 1; tpr[length(tpr)] <- 1
  list(fpr = fpr, tpr = tpr,
       auc = sum(diff(fpr) * (tpr[-1] + tpr[-length(tpr)]) / 2))
}

cols_roc <- c("steelblue3", "darkorange", "firebrick")

roc_list <- lapply(y_bin, function(y) {
  roc_points(cmp_b[[y]], cmp_b[[paste0("predicted_", y)]])
})
auc_vals <- sapply(roc_list, `[[`, "auc")

par(mfrow = c(1, 1), mar = c(4, 4, 3, 1))
plot(NULL, xlim = c(0, 1), ylim = c(0, 1),
     xlab = "False positive rate", ylab = "True positive rate",
     main = "ROC curves — binary VAR")
abline(0, 1, col = "grey70", lty = 2)
for (i in seq_along(y_bin)) {
  lines(roc_list[[i]]$fpr, roc_list[[i]]$tpr, col = cols_roc[i], lwd = 2)
}
legend("bottomright",
       legend = paste0(sub("_elev", "", y_bin),
                       " (AUC = ", round(auc_vals, 2), ")"),
       col = cols_roc, lwd = 2, bty = "n")
```

### 3c — Predicted probability trajectory

For a single participant we plot the day-by-day predicted probability of an
elevated day alongside the observed binary outcome.

```{r binary-trajectory, eval = has_cmdstan, fig.height = 4}
sd_b   <- tapply(fried_bin$alone_elev, fried_bin$id, sd, na.rm = TRUE)
subj_b <- names(which.max(sd_b))

par(mfrow = c(1, length(y_bin)), mar = c(4, 4, 3, 1))
for (i in seq_along(y_bin)) {
  y       <- y_bin[i]
  cmp_sub <- merge(
    fried_bin[fried_bin$id == subj_b, c("id", "day", y)],
    pred_bin[pred_bin$id == subj_b, c("id", "day", paste0("predicted_", y))],
    by = c("id", "day")
  )
  cmp_sub <- cmp_sub[order(cmp_sub$day), ]

  plot(cmp_sub$day, cmp_sub[[paste0("predicted_", y)]],
       type = "l", col = cols_roc[i], lwd = 2,
       ylim = c(0, 1), xlab = "Day", ylab = "P(elevated = 1)",
       main = sub("_elev", "", y))
  points(cmp_sub$day, jitter(cmp_sub[[y]], amount = 0.03),
         pch = 16, cex = 1.3, col = "black")
  abline(h = 0.5, col = "grey70", lty = 2)
}
```

The coloured line shows the predicted probability of an elevated day. Black
dots are the observed binary outcome (jittered for visibility). A dot near 1
on a day when the probability line is high indicates a correct prediction.

---

## References

Fried, E. I., Papanikolaou, F., & Epskamp, S. (2022). Mental health and social
contact during the COVID-19 pandemic: An ecological momentary assessment study.
*Clinical Psychological Science*, 10(2), 340–354.
<https://doi.org/10.1177/21677026211017839>

Wang, R., et al. (2020). StudentLife: Weakly supervised learning of mental
health using passively sensed campus behavioral data. In *Proceedings of the
ACM on Interactive, Mobile, Wearable and Ubiquitous Technologies*, 4(2), 1–26.
<https://doi.org/10.1145/3397321>

Siepe, B. S. (2025). openESM: Access the Open Experience Sampling Method
Database. R package. <https://openesmdata.org>

Blanken, T. F., et al. (2019). Introducing network intervention analysis to
investigate sequential, symptom-specific treatment effects. *JAMA Psychiatry*.
