---
title: "Prediction and Model Calibration (Simulated Data)"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Prediction and Model Calibration (Simulated Data)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse  = TRUE,
  comment   = "#>",
  fig.width = 7,
  fig.height = 5
)
has_cmdstan <- instantiate::stan_cmdstan_exists()
```

## Overview

This vignette mirrors `vignettes/Prediction_new.Rmd` but replaces real ESM
datasets with fully simulated data. Each dataset is generated from the **exact
probabilistic model** that `bvar()` fits, so the fitted predictions should
track the true generating process well. This serves two purposes:

1. **No external dependencies** — the vignette runs without internet access or
   cached dataset files.
2. **Ground-truth validation** — because we know the true parameters, we can
   check that the model recovers them and that the prediction plots look as
   expected for a well-specified model.

---

## Simulation backend

All simulated datasets in this vignette are generated with `bvarnet::sim_var()`,
which implements the **exact** generative model fit by `bvar()` for each
family (bernoulli, ordinal, gaussian). Using the package's own simulator
guarantees that what `bvar()` recovers matches what `sim_var()` puts in.

```{r sim-helpers}
library(bvarnet)

# Helper: rename sim_var output columns into the names used downstream.
# sim_var produces `t` (time) and `y_1..y_p` (outcomes); we rename to
# `day` and to user-supplied outcome names.
.relabel_sim <- function(sim_data, y_names) {
  names(sim_data)[names(sim_data) == "t"] <- "day"
  old <- paste0("y_", seq_along(y_names))
  names(sim_data)[match(old, names(sim_data))] <- y_names
  sim_data$id <- as.character(sim_data$id)
  sim_data
}
```


---

## 1 — Ordinal VAR

### True parameters

| Parameter | Value | Meaning |
|-----------|-------|---------|
| `phi` (diagonal) | 0.15 | Lag-1 autocorrelation on the latent (raw-lag) scale |
| `phi` (off-diagonal) | 0.05 | Cross-lagged coupling |
| `kappa` | (−2.5, −0.8, 0.8, 2.5) | Shared cutpoints; roughly bell-shaped marginal |
| σ_α | 0.40 | Between-subject random-intercept SD (on the latent scale) |

`sim_var(family = "ordinal")` uses the same parameterisation as the Stan
model: zero population intercept (absorbed into kappa), random intercepts
on the latent scale, and lag-1 effects applied to the **raw** previous
score. The φ values estimated by `bvar()` are therefore directly comparable
to the true values above.

```{r sim-ord-data, eval = has_cmdstan}
J <- 80L;  Tpp <- 50L;  p_ord <- 3L

phi_ord   <- matrix(0.05, p_ord, p_ord); diag(phi_ord) <- 0.15
kappa_ord <- list(c(-2.5, -0.8, 0.8, 2.5),
                  c(-2.5, -0.8, 0.8, 2.5),
                  c(-2.5, -0.8, 0.8, 2.5))
sd_u_ord  <- 0.10

sim_o   <- sim_var(N = J, T_obs = Tpp, p = p_ord, K = 1L,
                   family = "ordinal", Phi = phi_ord, kappa = kappa_ord,
                   sd_alpha = sd_u_ord, C = 5L, seed = 1337)
y_ord   <- paste0("item", seq_len(p_ord))
dat_ord <- .relabel_sim(sim_o$data, y_ord)

cat("Dimensions:", nrow(dat_ord), "×", ncol(dat_ord), "\n")
cat("Subjects:", length(unique(dat_ord$id)), "| Days:", max(dat_ord$day), "\n\n")
cat("Category frequencies per item:\n")
for (y in y_ord) {
  tbl <- table(dat_ord[[y]])
  cat(sprintf("  %-8s  %s\n", y,
              paste(names(tbl), format(as.integer(tbl), width = 4), sep = "=", collapse = "  ")))
}
```

### Fit the model

```{r fit-ord, eval = has_cmdstan, message = FALSE, warning = FALSE}
fit_ord <- bvar(
  id_col      = "id",
  time_col    = "day",
  y_cols      = y_ord,
  x_cols      = NULL,
  re_cols     = "Intercept",
  re_temporal = FALSE,
  K           = 1L,
  data        = dat_ord,
  family      = "ordinal",
  iter        = 4000,
  warmup      = 1000,
  chains      = 4,
  cores       = 4,
  seed        = 1337,
  save_data   = TRUE
)
```

```{r summary-ord, eval = has_cmdstan}
summary(fit_ord)
```

### Predict category probabilities

```{r predict-ord, eval = has_cmdstan}
probs_ord <- predict(fit_ord, type = "probabilities")
# Named list: one data.frame per item, columns id / day / cat_1 … cat_5
head(probs_ord[["item1"]])

# Shared constants used throughout the ordinal section
n_cats   <- 5L
cat_cols <- paste0("cat_", seq_len(n_cats))
ord_pal  <- hcl.colors(n_cats, "Blues 3")
obs_ord  <- fit_ord$data_used
```

### 1a — Calibration barplots

For each observed rating (x-axis), the stacked bar shows the mean predicted
probability across all categories. A well-calibrated model concentrates the
darkest bar (matching colour) at the correct column.

```{r calibration-ord, eval = has_cmdstan, fig.height = 5}
par(mfrow = c(1, p_ord), mar = c(4, 4, 2.5, 1))
for (y in y_ord) {
  cmp <- merge(obs_ord[, c("id", "day", y)],
               probs_ord[[y]], by = c("id", "day"))
  cal <- matrix(NA_real_, n_cats, n_cats)
  for (k in seq_len(n_cats)) {
    rows_k <- cmp[[y]] == k
    if (any(rows_k))
      cal[k, ] <- colMeans(cmp[rows_k, cat_cols, drop = FALSE])
  }
  barplot(t(cal), beside = FALSE, col = ord_pal,
          xlab = "Observed category", ylab = "Mean predicted prob",
          main = y, ylim = c(0, 1))
  legend("topright", legend = rev(seq_len(n_cats)), title = "Pred. cat.",
         fill = rev(ord_pal), cex = 0.65, bty = "n")
}
```

Because the data were generated from this exact model, the dominant bar in each
column should match the column index — any deviation reflects residual
estimation uncertainty.

### 1b — Individual probability trajectory

We pick the subject with the highest within-person SD in `item1` to show a
trajectory with clear category transitions.

```{r traj-ord, eval = has_cmdstan, fig.height = 5}
sd_per_subj <- tapply(obs_ord$item1, obs_ord$id, sd, na.rm = TRUE)
subj_pick   <- names(which.max(sd_per_subj))

cmp <- merge(
  obs_ord[obs_ord$id == subj_pick, c("id", "day", "item1")],
  probs_ord[["item1"]][probs_ord[["item1"]]$id == subj_pick, ],
  by = c("id", "day")
)
cmp <- cmp[order(cmp$day), ]

plot(NULL, xlim = range(cmp$day), ylim = c(0, 1),
     xlab = "Day", ylab = "Predicted probability",
     main = paste0("item1 — subject ", subj_pick))
cum <- numeric(nrow(cmp))
for (k in seq_len(n_cats)) {
  new_cum <- cum + cmp[[paste0("cat_", k)]]
  polygon(c(cmp$day, rev(cmp$day)),
          c(new_cum, rev(cum)), col = ord_pal[k], border = NA)
  cum <- new_cum
}
obs_scaled <- (cmp$item1 - 1) / (n_cats - 1)
points(cmp$day, obs_scaled, pch = 16, cex = 1.2)
legend("topright", legend = rev(seq_len(n_cats)), title = "Category",
       fill = rev(ord_pal), cex = 0.7, bty = "n")
```

Black dots show the (rescaled) observed category; a dot inside its matching
colour band indicates a correct modal prediction.

### 1c — Observed vs predicted frequency over time

Solid lines show observed category proportions aggregated over all subjects;
dashed lines show the model's mean predicted probability at the same time
points. Close tracking confirms the model captures population-level temporal
dynamics.

```{r freq-over-time-ord, eval = has_cmdstan, fig.height = 5}
par(mfrow = c(1, p_ord), mar = c(4, 4, 2.5, 1))
for (y in y_ord) {
  cmp  <- merge(obs_ord[, c("id", "day", y)], probs_ord[[y]], by = c("id", "day"))
  days <- sort(unique(cmp$day))
  obs_m  <- matrix(NA_real_, length(days), n_cats)
  pred_m <- matrix(NA_real_, length(days), n_cats)
  for (i in seq_along(days)) {
    rows_d   <- cmp$day == days[i]
    obs_cats <- cmp[[y]][rows_d]
    for (k in seq_len(n_cats)) {
      obs_m[i, k]  <- mean(obs_cats == k, na.rm = TRUE)
      pred_m[i, k] <- mean(cmp[[paste0("cat_", k)]][rows_d], na.rm = TRUE)
    }
  }
  ymax <- max(obs_m, pred_m, na.rm = TRUE)
  plot(NULL, xlim = range(days), ylim = c(0, ymax),
       xlab = "Day", ylab = "Frequency / probability", main = y)
  for (k in seq_len(n_cats)) {
    lines(days, obs_m[, k],  col = ord_pal[k], lwd = 2, lty = 1)
    lines(days, pred_m[, k], col = ord_pal[k], lwd = 2, lty = 2)
  }
  legend("topright",
         legend = c(paste0("Cat. ", seq_len(n_cats)), "Obs.", "Pred."),
         col    = c(ord_pal, "black", "black"),
         lty    = c(rep(NA, n_cats), 1L, 2L),
         pch    = c(rep(15L, n_cats), NA, NA),
         lwd = 2, cex = 0.55, bty = "n", ncol = 2)
}
```

### 1d — Modal category accuracy

The table reports the fraction of observations for which the model's modal
predicted category matches the observed value. The majority-class baseline
is the accuracy of always predicting the most frequent observed category —
a more meaningful reference than uniform chance (1/5 = 20%) for skewed
distributions. Model accuracy should exceed this baseline to demonstrate
that the lag structure adds genuine predictive value.

```{r modal-acc-ord, eval = has_cmdstan}
acc <- data.frame(variable = y_ord, pct_correct = NA_real_, pct_majority = NA_real_)
for (i in seq_along(y_ord)) {
  y   <- y_ord[i]
  cmp <- merge(obs_ord[, c("id", "day", y)],
               probs_ord[[y]], by = c("id", "day"))
  modal_pred  <- max.col(cmp[, cat_cols])
  acc$pct_correct[i]  <- round(100 * mean(modal_pred == cmp[[y]], na.rm = TRUE), 1)
  # Majority-class baseline: always predict the most frequent observed category
  majority_cat <- as.integer(names(which.max(table(cmp[[y]]))))
  acc$pct_majority[i] <- round(100 * mean(majority_cat == cmp[[y]], na.rm = TRUE), 1)
}
names(acc) <- c("Variable", "% correct (model)", "% correct (majority baseline)")
knitr::kable(acc, align = "lrr")
```

---

## 2 — Gaussian VAR

### True parameters

| Parameter | Value | Meaning |
|-----------|-------|---------|
| `phi` (diagonal) | 0.50, 0.40 | Lag-1 autocorrelation per outcome |
| `phi` (off-diagonal) | 0.05 | Cross-lagged coupling |
| `beta` | 0, 0 | Group mean of each outcome |
| σ | 0.70 | Residual (within-person) SD |
| σ_u | 1.00 | Between-subject SD for random intercepts |

With σ_u = 1 and σ = 0.7, between-person variation dominates, which is typical
for psychological ESM outcomes. The one-step prediction R² at the within-person
level is approximately φ² ≈ 0.25 / 0.16 for each outcome.

```{r sim-gauss-data, eval = has_cmdstan}
p_gauss     <- 2L
phi_gauss   <- matrix(0.05, p_gauss, p_gauss); diag(phi_gauss) <- c(0.50, 0.40)
alpha_gauss <- rep(0, p_gauss)
sigma_gauss <- rep(0.10, p_gauss)
sd_u_gauss  <- 0.15

sim_g     <- sim_var(N = J, T_obs = Tpp, p = p_gauss, K = 1L,
                     family = "gaussian", Phi = phi_gauss,
                     alpha = alpha_gauss, sigma = sigma_gauss,
                     sd_alpha = sd_u_gauss, seed = 1337)
y_gauss   <- paste0("y", seq_len(p_gauss))
dat_gauss <- .relabel_sim(sim_g$data, y_gauss)

cat("Dimensions:", nrow(dat_gauss), "×", ncol(dat_gauss), "\n")
for (y in y_gauss) {
  cat(sprintf("  %-4s  M = %5.2f  SD = %5.2f  range [%5.2f, %5.2f]\n",
              y,
              mean(dat_gauss[[y]], na.rm = TRUE),
              sd(dat_gauss[[y]], na.rm = TRUE),
              min(dat_gauss[[y]], na.rm = TRUE),
              max(dat_gauss[[y]], na.rm = TRUE)))
}
```

### Fit the model

```{r fit-gauss, eval = has_cmdstan, message = FALSE, warning = FALSE}
fit_gauss <- bvar(
  id_col      = "id",
  time_col    = "day",
  y_cols      = y_gauss,
  x_cols      = NULL,
  re_cols     = "Intercept",
  re_temporal = FALSE,
  K           = 1L,
  data        = dat_gauss,
  family      = "gaussian",
  iter        = 4000,
  warmup      = 1000,
  chains      = 4,
  cores       = 4,
  seed        = 1337,
  save_data   = TRUE
)
```

```{r summary-gauss, eval = has_cmdstan}
summary(fit_gauss)
```

### 2a — Observed vs predicted scatter

For Gaussian outcomes the fitted values are on the same scale as the data. We
also annotate the true φ (diagonal) so the reader can see how well the model
recovers the generative parameter.

```{r scatter-gauss, eval = has_cmdstan, fig.height = 5}
pred_gauss <- predict(fit_gauss, type = "response")
obs_gauss  <- fit_gauss$data_used
cmp_g      <- merge(obs_gauss, pred_gauss, by = c("id", "day"))

true_phi_d <- diag(phi_gauss)
par(mfrow = c(1, p_gauss), mar = c(4, 4, 3, 1))
for (i in seq_len(p_gauss)) {
  y        <- y_gauss[i]
  obs_v    <- cmp_g[[y]]
  pred_v   <- cmp_g[[paste0("predicted_", y)]]
  r2       <- cor(obs_v, pred_v, use = "complete.obs")^2

  plot(obs_v, pred_v, pch = 16, cex = 0.5,
       col  = rgb(0.2, 0.4, 0.8, 0.35),
       xlab = paste("Observed", y),
       ylab = paste("Predicted", y),
       main = paste0(y, "  (true φ = ", true_phi_d[i], ",  R² = ",
                     round(r2, 2), ")"))
  abline(0, 1, col = "firebrick", lwd = 2)
}
```

The identity line (red) shows perfect one-step prediction. Points lie away from
it primarily because the total variance includes between-person differences
(σ_u = 1) that are captured by the random intercept, not the lag coefficient.

### 2b — Residual distribution

```{r resid-gauss, eval = has_cmdstan, fig.height = 4}
par(mfrow = c(1, p_gauss), mar = c(4, 4, 3, 1))
for (y in y_gauss) {
  resid <- cmp_g[[y]] - cmp_g[[paste0("predicted_", y)]]
  hist(resid, breaks = 35,
       xlab = "Residual (observed − predicted)",
       main = paste0(y, " residuals"),
       col  = "steelblue3", border = "white", freq = FALSE)
  curve(dnorm(x, 0, sd(resid, na.rm = TRUE)),
        col = "firebrick", lwd = 2, add = TRUE)
}
```

The residuals should be approximately Gaussian with mean zero (Gaussian family
assumption). Because the data were generated with the same assumption, strong
non-normality here would indicate a fitting problem rather than a
misspecification.

---

## 3 — Binary VAR

### True parameters

| Parameter | Value | Meaning |
|-----------|-------|---------|
| `phi` (diagonal) | 0.50 | Lag-1 autocorrelation on logit scale |
| `phi` (off-diagonal) | 0.10 | Cross-lagged coupling (logit scale) |
| `alpha` | −0.50 | Intercept → base rate ≈ 38% at u = 0 |
| σ_α | 0.50 | Between-subject SD (logit scale) |

A φ of 0.50 on the logit scale means that a previous positive event shifts the
log-odds by +0.50, raising the predicted probability by roughly 10
percentage points near the base rate. This produces AUC above 0.5 and a
clear reliability-diagram signal while keeping the VAR system stable.

```{r sim-bin-data, eval = has_cmdstan}
p_bin     <- 3L
phi_bin   <- matrix(0.10, p_bin, p_bin); diag(phi_bin) <- 0.50
alpha_bin <- rep(-0.50, p_bin)
sd_u_bin  <- 0.10

sim_b   <- sim_var(N = J, T_obs = Tpp, p = p_bin, K = 1L,
                   family = "bernoulli", Phi = phi_bin,
                   alpha = alpha_bin, sd_alpha = sd_u_bin, seed = 1337)
y_bin   <- paste0("event", seq_len(p_bin))
dat_bin <- .relabel_sim(sim_b$data, y_bin)

cat("Base rates:\n")
for (y in y_bin)
  cat(sprintf("  %-8s  %s\n", y,
              paste0(round(100 * mean(dat_bin[[y]], na.rm = TRUE), 1), "%")))
```

### Fit the model

```{r fit-bin, eval = has_cmdstan, message = FALSE, warning = FALSE}
fit_bin <- bvar(
  id_col      = "id",
  time_col    = "day",
  y_cols      = y_bin,
  x_cols      = NULL,
  re_cols     = "Intercept",
  re_temporal = FALSE,
  K           = 1L,
  data        = dat_bin,
  family      = "bernoulli",
  iter        = 4000,
  warmup      = 1000,
  chains      = 4,
  cores       = 4,
  seed        = 1337,
  save_data   = TRUE
)
```

```{r summary-bin, eval = has_cmdstan}
summary(fit_bin)
```

### 3a — Observed vs predicted event frequency over time

For each binary outcome, we aggregate across subjects at every day and compare
the **observed event rate** (dotted line) with the **mean predicted
probability** (solid line), in the style of Schumacher et al. (2023,
*JAMA Psychiatry*, eFigure 1). Per-observation predicted probabilities are
plotted as faint dots in the background so the day-level summaries can be
read against the spread of individual-level predictions.

```{r freq-bin, eval = has_cmdstan, fig.height = 4}
pred_bin <- predict(fit_bin, type = "response")
obs_bin  <- fit_bin$data_used
cmp_b    <- merge(obs_bin, pred_bin, by = c("id", "day"))

cols_roc <- c("steelblue3", "darkorange", "firebrick")

days <- sort(unique(cmp_b$day))
obs_m  <- matrix(NA_real_, length(days), p_bin)
pred_m <- matrix(NA_real_, length(days), p_bin)
for (i in seq_along(days)) {
  rows_d <- cmp_b$day == days[i]
  for (j in seq_len(p_bin)) {
    obs_m[i, j]  <- mean(cmp_b[[y_bin[j]]][rows_d], na.rm = TRUE)
    pred_m[i, j] <- mean(cmp_b[[paste0("predicted_", y_bin[j])]][rows_d],
                         na.rm = TRUE)
  }
}

par(mfrow = c(1, p_bin), mar = c(4, 4, 3, 1))
for (j in seq_len(p_bin)) {
  y    <- y_bin[j]
  plot(NULL, xlim = range(days), ylim = c(0, 1),
       xlab = "Day", ylab = "P(event = 1)",
       main = y)
  # individual predicted probabilities as background dots
  points(jitter(cmp_b$day, amount = 0.25),
         cmp_b[[paste0("predicted_", y)]],
         pch = 16, cex = 0.35,
         col = adjustcolor(cols_roc[j], alpha.f = 0.18))
  # observed event rate (dotted) and mean predicted probability (solid)
  lines(days, obs_m[, j],  col = cols_roc[j], lwd = 2, lty = 3)
  lines(days, pred_m[, j], col = cols_roc[j], lwd = 2, lty = 1)
  legend("topright",
         legend = c("Observed (rate)", "Predicted (mean prob.)",
                    "Per-obs predicted prob."),
         col    = c(cols_roc[j], cols_roc[j],
                    adjustcolor(cols_roc[j], alpha.f = 0.4)),
         lty    = c(3, 1, NA),
         pch    = c(NA, NA, 16),
         lwd    = c(2, 2, NA),
         cex = 0.65, bty = "n")
}
```

If the solid (predicted) line closely tracks the dotted (observed) line, the
model is calibrated **at the day level**: it gets the right average
prevalence on each day, even if it is uncertain about which specific
occasions the event occurs on.

### 3b — Predicted probability trajectory

For the subject with the most variable `event1`, we plot the day-by-day
predicted P(event = 1) alongside the observed binary outcome.

```{r traj-bin, eval = has_cmdstan, fig.height = 4}
sd_b   <- tapply(dat_bin$event1, dat_bin$id, sd, na.rm = TRUE)
subj_b <- names(which.max(sd_b))

par(mfrow = c(1, p_bin), mar = c(4, 4, 3, 1))
for (i in seq_len(p_bin)) {
  y <- y_bin[i]
  sub_obs  <- obs_bin[obs_bin$id == subj_b, c("id", "day", y)]
  sub_pred <- pred_bin[pred_bin$id == subj_b, c("id", "day", paste0("predicted_", y))]
  cmp_sub  <- merge(sub_obs, sub_pred, by = c("id", "day"))
  cmp_sub  <- cmp_sub[order(cmp_sub$day), ]

  plot(cmp_sub$day, cmp_sub[[paste0("predicted_", y)]],
       type = "l", col = cols_roc[i], lwd = 2,
       ylim = c(0, 1), xlab = "Day",
       ylab = "P(event = 1)", main = paste0(y, " — subject ", subj_b))
  points(cmp_sub$day,
         jitter(cmp_sub[[y]], amount = 0.03),
         pch = 16, cex = 1.3)
  abline(h = 0.5, col = "grey70", lty = 2)
}
```

The coloured line is the model's predicted probability; black dots are the
observed outcome (1 = event occurred, 0 = not; jittered by 0.03 for
visibility). Days where the probability line crosses 0.5 correspond to the
predicted category boundary.

### 3c — Recursive forecast from a single starting point

Sections 3a–3b use **one-step-ahead, in-sample** predictions: every
prediction conditions on the *observed* lag-1 value at that day. This is
why the solid (predicted) line in §3a looks much smoother than the dotted
(observed) line — it is a conditional expectation $\Pr(Y_t = 1 \mid Y_{t-1})$,
re-anchored at every step by the true data.

A more demanding question is **forecast skill**: given only the first few
days as a conditioning window, how well does the model project the
trajectory forward when its own predicted values are fed back into the
lag buffer? We use `forecast = "recursive"` with a 5-day conditioning window
and `method = "posterior-sample"` so we can show a 95% band across draws.

```{r recursive-bin, eval = has_cmdstan, fig.height = 4}
cw <- 5L  # condition on the first 5 days, recursively forecast the rest

pred_rec <- predict(
  fit_bin, newdata = dat_bin, type = "response",
  forecast = "recursive", conditioning_window = cw,
  method   = "posterior-sample", ndraws = 200, seed = 1337
)

cmp_r <- merge(obs_bin, pred_rec, by = c("id", "day"))

days  <- sort(unique(cmp_r$day))
obs_r <- pred_r <- lo_r <- hi_r <- matrix(NA_real_, length(days), p_bin)
for (i in seq_along(days)) {
  rows_d <- cmp_r$day == days[i]
  for (j in seq_len(p_bin)) {
    y    <- y_bin[j]
    obs_r[i, j]  <- mean(cmp_r[[y]][rows_d], na.rm = TRUE)
    mu_d <- cmp_r[[paste0("predicted_", y)]][rows_d]
    sd_d <- cmp_r[[paste0("predicted_", y, "_sd")]][rows_d]
    pred_r[i, j] <- mean(mu_d, na.rm = TRUE)
    # day-level 95% band on the mean predicted rate (across-draw uncertainty
    # of the population average): SD of the mean ≈ mean(sd) / sqrt(n_per_day)
    n_d <- sum(!is.na(mu_d))
    se  <- if (n_d > 0) mean(sd_d, na.rm = TRUE) / sqrt(n_d) else NA_real_
    lo_r[i, j] <- pmax(0, pred_r[i, j] - 1.96 * se)
    hi_r[i, j] <- pmin(1, pred_r[i, j] + 1.96 * se)
  }
}

par(mfrow = c(1, p_bin), mar = c(4, 4, 3, 1))
for (j in seq_len(p_bin)) {
  y <- y_bin[j]
  plot(NULL, xlim = range(days), ylim = c(0, 1),
       xlab = "Day", ylab = "P(event = 1)", main = y)
  # shade the conditioning window
  rect(min(days) - 0.5, 0, min(days) + cw - 0.5, 1,
       col = adjustcolor("grey80", alpha.f = 0.4), border = NA)
  # 95% band on the recursive predicted mean
  polygon(c(days, rev(days)), c(lo_r[, j], rev(hi_r[, j])),
          col = adjustcolor(cols_roc[j], alpha.f = 0.20), border = NA)
  lines(days, obs_r[,  j], col = cols_roc[j], lwd = 2, lty = 3)
  lines(days, pred_r[, j], col = cols_roc[j], lwd = 2, lty = 1)
  abline(v = min(days) + cw - 0.5, col = "grey50", lty = 2)
  legend("topright",
         legend = c("Observed (rate)", "Recursive forecast",
                    "95% band", "Conditioning window"),
         col    = c(cols_roc[j], cols_roc[j],
                    adjustcolor(cols_roc[j], alpha.f = 0.4),
                    adjustcolor("grey80", alpha.f = 0.6)),
         lty = c(3, 1, NA, NA), pch = c(NA, NA, 15, 15),
         lwd = c(2, 2, NA, NA), cex = 0.6, bty = "n")
}
```

Two features distinguish this from §3d:

1. **The forecast line is no longer pinned to the data each step.** After
   day `r I("cw")`, the model uses *its own* predicted probabilities to
   update the lag buffer. With true φ around 0.5, the recursive forecast
   relaxes toward each subject's stationary mean over a few days, then
   stays there.
2. **The forecast line is much flatter than the observed-rate line.**
   Because the only signal driving day-to-day variation in the recursive
   forecast is the propagated lag, and that lag has decayed toward the
   stationary mean, the population-average forecast smooths out the
   day-level noise that the one-step-ahead prediction was tracking. The
   observed daily rates therefore wobble around — but no longer through —
   the recursive forecast curve. This is the expected pattern: **good
   in-sample calibration does not imply good multi-step forecast skill**,
   and vice versa.
