---
title: "Prediction and Model Calibration"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Prediction and Model Calibration}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse  = TRUE,
  comment   = "#>",
  fig.width = 7,
  fig.height = 5
)
has_cmdstan <- instantiate::stan_cmdstan_exists()
```

## Overview

The `predict()` method for **bvarnet** objects generates fitted values for the
two response families covered in this vignette:

| Family | `type = "link"` | `type = "response"` | `type = "probabilities"` |
|--------|----------------|---------------------|--------------------------|
| `"ordinal"`  | latent linear predictor | modal category | category probabilities (list of data frames) |
| `"gaussian"` | fitted mean (= response scale) | fitted mean | — |

We illustrate prediction with two openly available ESM datasets accessed via
the `openesm` package (Siepe 2025). `openesm::get_dataset()` serves cached
data on subsequent calls.

**Dataset selection criterion.** Multilevel VAR prediction plots are only
informative if the outcomes carry genuine *within-person* temporal signal —
i.e., if knowing yesterday's value (after removing each person's mean) helps
predict today's. After screening 15 cached openESM datasets on within-person
lag-1 autocorrelation, distribution shape, and coverage, we use:

* **Geschwind et al. (2013)** — 130 individuals with depression in an MBCT
  trial, 7-point momentary affect items, ~220 beeps per subject. Within-person
  lag-1 r ≈ 0.35 – 0.42.
* **Fisher et al. (2019)** — 52 regular smokers, 0–100 visual-analog
  motivation/craving sliders, ~124 beeps per subject. Within-person lag-1
  r ≈ 0.21 – 0.23.

We deliberately omit a **binary** section: across all surveyed openESM
datasets, no clinically meaningful binary outcome had both adequate base
rates and non-trivial autocorrelation, and binarising ordinal items
destroys most of their within-person signal (e.g. `sad ≥ 3` drops from
lag-1 ≈ 0.42 to ≈ 0.12). The companion vignette `Predict_sim.Rmd`
demonstrates the binary workflow on simulated data.

---

## 1 — Ordinal VAR: Momentary Affect in Depression (Geschwind 2013)

The Geschwind (2013) dataset comes from a randomised controlled trial of
mindfulness-based cognitive therapy in 130 patients with residual depressive
symptoms. Participants received roughly 10 random prompts per day for 6 days
before and after treatment, rating momentary affect on a 7-point scale (1 =
*not at all*, 7 = *very*). We model four core affect items with strong
within-person dynamics:

| Item       | Lag-1 *r* | Modal share |
|------------|-----------|-------------|
| `cheerful` | 0.42      | 0.27 |
| `sad`      | 0.42      | 0.45 |
| `worried`  | 0.41      | 0.39 |
| `relaxed`  | 0.35      | 0.27 |

```{r load-ordinal, eval = has_cmdstan, message = FALSE, warning = FALSE}
library(bvarnet)
library(openesm)

geschwind_raw <- get_dataset("0010_geschwind")$data

y_ord <- c("cheerful", "sad", "worried", "relaxed")

geschwind <- geschwind_raw[complete.cases(geschwind_raw[, c("id", "day", "beep", y_ord)]), ]
geschwind$id <- as.character(geschwind$id)

# bvar() requires a unique time index per (subject, time). Geschwind has
# multiple beeps per day, so we build a sequential within-subject time stamp.
geschwind <- geschwind[order(geschwind$id, geschwind$day, geschwind$beep), ]
geschwind$t <- ave(seq_len(nrow(geschwind)),
                   geschwind$id, FUN = seq_along)

cat("Subjects:", length(unique(geschwind$id)),
    "| Total observations:", nrow(geschwind), "\n")
cat("Median beeps per subject:",
    median(as.integer(table(geschwind$id))), "\n")
cat("Response scale: 1 – 7 (not at all … very)\n")
```

### Fit the model

```{r fit-ordinal, eval = has_cmdstan, message = FALSE, warning = FALSE}
fit_ord <- bvar(
  id_col      = "id",
  time_col    = "t",
  y_cols      = y_ord,
  x_cols      = NULL,
  re_cols     = c("Intercept"),
  re_temporal = FALSE,
  K           = 1,
  data        = geschwind,
  family      = "ordinal",
  iter        = 4000,
  warmup      = 1000,
  chains      = 4,
  cores       = 4,
  seed        = 42,
  save_data   = TRUE
)
```

```{r summary-ordinal, eval = has_cmdstan}
summary(fit_ord)
```

### 1a — What does `predict()` return for ordinal models?

The default call returns one-step-ahead fitted values on the **latent linear
predictor** scale. For ordinal models these are not directly interpretable, so
we move straight to category probabilities.

```{r predict-probs-ordinal, eval = has_cmdstan}
probs_ord <- predict(fit_ord, type = "probabilities")
# probs_ord is a named list: one data.frame per outcome
# Each data.frame has columns id, t, cat_1 … cat_7
head(probs_ord[["sad"]])

n_cats   <- 7L
cat_cols <- paste0("cat_", seq_len(n_cats))
ord_pal  <- hcl.colors(n_cats, "Blues 3")
```

### 1b — Calibration: are the right categories predicted?

A well-calibrated model assigns the **highest probability to the observed
category**. The barplot shows, for each observed rating (x-axis), the mean
predicted probability per category (stacked bars). A diagonal pattern — dark
blue dominating low ratings and shifting to light blue at high ratings —
indicates the model tracks ordinal responses.

```{r calibration-barplot, eval = has_cmdstan, fig.height = 7}
obs_ord <- fit_ord$data_used

par(mfrow = c(2, 2), mar = c(4, 4, 2.5, 1))
for (y in y_ord) {
  cmp <- merge(
    obs_ord[, c("id", "t", y)],
    probs_ord[[y]],
    by = c("id", "t")
  )
  cal <- matrix(NA_real_, n_cats, n_cats,
                dimnames = list(observed  = seq_len(n_cats),
                                predicted = cat_cols))
  for (k in seq_len(n_cats)) {
    rows_k <- cmp[[y]] == k
    if (any(rows_k)) cal[k, ] <- colMeans(cmp[rows_k, cat_cols, drop = FALSE])
  }
  barplot(
    t(cal), beside = FALSE, col = ord_pal,
    xlab = "Observed category", ylab = "Mean predicted probability",
    main = y, ylim = c(0, 1)
  )
  legend("topright",
         legend = rev(seq_len(n_cats)), title = "Predicted cat.",
         fill = rev(ord_pal), cex = 0.6, bty = "n")
}
```

### 1c — Individual probability trajectory

The stacked area chart shows how predicted category probabilities evolve across
beeps for the participant with the highest within-person variability in `sad`.
Black dots mark the rescaled observed rating (1 → 0, 7 → 1); a dot inside its
corresponding coloured band indicates a correct modal prediction.

```{r trajectory, eval = has_cmdstan, fig.height = 5}
sd_per_subj <- tapply(obs_ord$sad, obs_ord$id, sd, na.rm = TRUE)
subj_pick   <- names(which.max(sd_per_subj))

y <- "sad"
cmp <- merge(
  obs_ord[obs_ord$id == subj_pick, c("id", "t", y)],
  probs_ord[[y]][probs_ord[[y]]$id == subj_pick, ],
  by = c("id", "t")
)
cmp <- cmp[order(cmp$t), ]

plot(NULL, xlim = range(cmp$t), ylim = c(0, 1),
     xlab = "Beep (within subject)", ylab = "Predicted probability",
     main = paste0(y, " — participant ", subj_pick))

cum <- numeric(nrow(cmp))
for (k in seq_len(n_cats)) {
  new_cum <- cum + cmp[[paste0("cat_", k)]]
  polygon(c(cmp$t, rev(cmp$t)), c(new_cum, rev(cum)),
          col = ord_pal[k], border = NA)
  cum <- new_cum
}
points(cmp$t, (cmp[[y]] - 1) / (n_cats - 1), pch = 16, cex = 1.0)
legend("topright", legend = rev(seq_len(n_cats)), title = "Category",
       fill = rev(ord_pal), cex = 0.6, bty = "n")
```

### 1d — Observed vs predicted category frequencies over time

We compare observed category proportions (solid lines) with mean predicted
probabilities (dashed) aggregated across all subjects at each within-subject
time index. Close tracking indicates the model captures population-level
temporal dynamics. To keep the panels readable we bin the time index into
20 equal-width windows.

```{r freq-over-time, eval = has_cmdstan, fig.height = 7}
n_bins_t <- 20

par(mfrow = c(2, 2), mar = c(4, 4, 2.5, 1))
for (y in y_ord) {
  cmp <- merge(obs_ord[, c("id", "t", y)], probs_ord[[y]], by = c("id", "t"))
  cmp$tbin <- cut(cmp$t, breaks = n_bins_t, labels = FALSE,
                  include.lowest = TRUE)
  bins <- sort(unique(cmp$tbin))
  bin_mid <- tapply(cmp$t, cmp$tbin, mean, na.rm = TRUE)

  obs_m  <- matrix(NA_real_, length(bins), n_cats)
  pred_m <- matrix(NA_real_, length(bins), n_cats)
  for (i in seq_along(bins)) {
    rows_b   <- cmp$tbin == bins[i]
    obs_cats <- cmp[[y]][rows_b]
    for (k in seq_len(n_cats)) {
      obs_m[i, k]  <- mean(obs_cats == k, na.rm = TRUE)
      pred_m[i, k] <- mean(cmp[[paste0("cat_", k)]][rows_b], na.rm = TRUE)
    }
  }
  ymax <- max(obs_m, pred_m, na.rm = TRUE)
  plot(NULL, xlim = range(bin_mid), ylim = c(0, ymax),
       xlab = "Beep (binned)", ylab = "Proportion", main = y)
  for (k in seq_len(n_cats)) {
    lines(bin_mid, obs_m[, k],  col = ord_pal[k], lwd = 2, lty = 1)
    lines(bin_mid, pred_m[, k], col = ord_pal[k], lwd = 2, lty = 2)
  }
  legend("topright",
         legend = c(paste0("Cat. ", seq_len(n_cats)), "Observed", "Predicted"),
         col    = c(ord_pal, "black", "black"),
         lty    = c(rep(NA, n_cats), 1, 2),
         pch    = c(rep(15, n_cats), NA, NA),
         lwd = 2, cex = 0.5, bty = "n", ncol = 2)
}
```

### 1e — Modal category accuracy

The proportion of observations for which the predicted modal category matches
the observed response, alongside a majority-class baseline (always predicting
the most common category).

```{r modal-accuracy, eval = has_cmdstan}
acc_tbl <- data.frame(
  variable     = y_ord,
  pct_correct  = NA_real_,
  pct_majority = NA_real_
)
for (i in seq_along(y_ord)) {
  y   <- y_ord[i]
  cmp <- merge(obs_ord[, c("id", "t", y)], probs_ord[[y]], by = c("id", "t"))
  modal_pred              <- max.col(cmp[, cat_cols])
  acc_tbl$pct_correct[i]  <- round(100 * mean(modal_pred == cmp[[y]], na.rm = TRUE), 1)
  majority_cat            <- as.integer(names(which.max(table(cmp[[y]]))))
  acc_tbl$pct_majority[i] <- round(100 * mean(majority_cat == cmp[[y]], na.rm = TRUE), 1)
}
names(acc_tbl) <- c("Variable", "% correct (model)", "% correct (majority baseline)")
knitr::kable(acc_tbl, align = "lrr")
```

---

## 2 — Gaussian VAR

This section illustrates two complementary Gaussian use cases:

* **2A**: the Geschwind 7-point items treated as approximately continuous —
  the Likert spacing is fine enough that a Gaussian approximation is often
  used in network/VAR analyses, and the strong within-person dynamics make for
  informative prediction plots.
* **2B**: the Fisher (2019) 0–100 visual-analog sliders — genuinely
  continuous outcomes from a different ESM design.

### 2A — Geschwind 7-point affect items as Gaussian

```{r fit-gaussian-geschwind, eval = has_cmdstan, message = FALSE, warning = FALSE}
fit_gauss_g <- bvar(
  id_col      = "id",
  time_col    = "t",
  y_cols      = y_ord,
  x_cols      = NULL,
  re_cols     = c("Intercept"),
  re_temporal = FALSE,
  K           = 1,
  data        = geschwind,
  family      = "gaussian",
  iter        = 4000,
  warmup      = 1000,
  chains      = 4,
  cores       = 4,
  seed        = 42,
  save_data   = TRUE
)
```

```{r summary-gaussian-geschwind, eval = has_cmdstan}
summary(fit_gauss_g)
```

#### Observed vs predicted

Points close to the identity line (red) indicate the model captures
within-person variation beyond each person's mean level. R² is the squared
Pearson correlation between observed and fitted values.

```{r gaussian-geschwind-scatter, eval = has_cmdstan, fig.height = 7}
pred_g  <- predict(fit_gauss_g, type = "response")
obs_g   <- fit_gauss_g$data_used
cmp_g   <- merge(obs_g, pred_g, by = c("id", "t"))

par(mfrow = c(2, 2), mar = c(4, 4, 2.5, 1))
for (y in y_ord) {
  observed  <- cmp_g[[y]]
  predicted <- cmp_g[[paste0("predicted_", y)]]
  r2        <- cor(observed, predicted, use = "complete.obs")^2
  plot(jitter(observed, amount = 0.2), predicted,
       xlab = paste("Observed", y), ylab = paste("Predicted", y),
       main = paste0(y, "  (R\u00b2 = ", round(r2, 2), ")"),
       pch = 16, cex = 0.4, col = rgb(0.2, 0.4, 0.8, 0.25))
  abline(0, 1, col = "firebrick", lwd = 2)
}
```

Observed values are jittered horizontally to reveal the density of points at
each integer rating. Predicted values are continuous because the Gaussian
likelihood does not enforce category boundaries — a reminder that treating a
7-point scale as Gaussian is an approximation.

#### Residuals

```{r gaussian-geschwind-residuals, eval = has_cmdstan, fig.height = 7}
par(mfrow = c(2, 2), mar = c(4, 4, 2.5, 1))
for (y in y_ord) {
  resid <- cmp_g[[y]] - cmp_g[[paste0("predicted_", y)]]
  hist(resid, breaks = 30, freq = FALSE,
       xlab = "Residual (observed \u2212 predicted)",
       main = paste0(y, " residuals"),
       col = "steelblue3", border = "white")
  curve(dnorm(x, mean = 0, sd = sd(resid, na.rm = TRUE)),
        col = "firebrick", lwd = 2, add = TRUE)
}
```

#### Prediction intervals

A point prediction is incomplete without an associated uncertainty band. We
request posterior-sample predictions (`method = "posterior-sample"`), which
returns the posterior-mean fitted value together with `_sd` columns capturing
uncertainty in the conditional mean across draws. A 95% **prediction
interval** for a future observation also has to absorb residual noise:

$$
\widehat y \;\pm\; 1.96 \,\sqrt{ \mathrm{SD}_{\mu}^{2} \;+\; \hat\sigma_{\varepsilon}^{2} },
$$

where $\mathrm{SD}_{\mu}$ is the across-draw SD of the conditional mean and
$\hat\sigma_{\varepsilon}$ is the posterior-mean residual SD per outcome.

```{r gaussian-geschwind-pi, eval = has_cmdstan, fig.height = 7}
pred_g_s <- predict(fit_gauss_g, type = "response",
                    method = "posterior-sample", ndraws = 200, seed = 42)

# Posterior-mean residual SD per outcome
ptab    <- extract_param(fit_gauss_g)
sigma_g <- setNames(ptab$mean[ptab$type == "sigma"],
                    ptab$outcome[ptab$type == "sigma"])

# Pick the subject with the most observations for visualisation
subj_pi <- names(which.max(table(pred_g_s$id)))
cmp_pi  <- merge(
  obs_g[obs_g$id == subj_pi, c("id", "t", y_ord)],
  pred_g_s[pred_g_s$id == subj_pi, ],
  by = c("id", "t")
)
cmp_pi <- cmp_pi[order(cmp_pi$t), ]

par(mfrow = c(2, 2), mar = c(4, 4, 2.5, 1))
for (y in y_ord) {
  mu_hat <- cmp_pi[[paste0("predicted_", y)]]
  sd_mu  <- cmp_pi[[paste0("predicted_", y, "_sd")]]
  pi_w   <- 1.96 * sqrt(sd_mu^2 + sigma_g[[y]]^2)
  lo     <- mu_hat - pi_w
  hi     <- mu_hat + pi_w

  plot(NULL, xlim = range(cmp_pi$t), ylim = range(c(lo, hi, cmp_pi[[y]]), na.rm = TRUE),
       xlab = "Beep (within subject)", ylab = y,
       main = paste0(y, " — participant ", subj_pi))
  polygon(c(cmp_pi$t, rev(cmp_pi$t)), c(lo, rev(hi)),
          col = rgb(0.2, 0.4, 0.8, 0.20), border = NA)
  lines(cmp_pi$t, mu_hat, col = "steelblue4", lwd = 2)
  points(cmp_pi$t, cmp_pi[[y]], pch = 16, cex = 0.8, col = "black")
  legend("topright",
         legend = c("Observed", "Predicted mean", "95% PI"),
         pch    = c(16, NA, 15),
         lty    = c(NA, 1, NA),
         col    = c("black", "steelblue4", rgb(0.2, 0.4, 0.8, 0.4)),
         lwd    = 2, cex = 0.7, bty = "n")

  cov95 <- mean(cmp_pi[[y]] >= lo & cmp_pi[[y]] <= hi, na.rm = TRUE)
  mtext(sprintf("empirical 95%% PI coverage: %.1f%%", 100 * cov95),
        side = 3, line = 0.1, cex = 0.7, col = "grey30")
}
```

For a well-specified Gaussian model the empirical coverage of the 95%
prediction interval should be close to 95%. Coverage well below this signals
residual heteroscedasticity, non-normality, or model misspecification — for
discretised 7-point data, some shortfall is expected because the Gaussian
likelihood ignores the bounded, integer support of the outcome.

### 2B — Fisher 2019 visual-analog sliders

Fisher (2019) sampled 52 regular smokers approximately 4 times per day over
roughly one month. Each prompt collected a battery of 0–100 sliders. We
model three motivation/craving items with the strongest within-person
autocorrelation in this dataset:

| Item                     | Lag-1 *r* | Range |
|--------------------------|-----------|-------|
| `motivated_quit_smoking` | 0.22      | 0–100 |
| `want_quit`              | 0.23      | 0–100 |
| `enjoy_last`             | 0.21      | 0–100 |

```{r load-fisher, eval = has_cmdstan, message = FALSE, warning = FALSE}
fisher_raw <- get_dataset("0037_fisher")$data

y_fisher <- c("motivated_quit_smoking", "want_quit", "enjoy_last")

fisher <- fisher_raw[complete.cases(fisher_raw[, c("id", "beep", y_fisher)]), ]
fisher$id <- as.character(fisher$id)

# Build a sequential within-subject time index
fisher <- fisher[order(fisher$id, fisher$beep), ]
fisher$t <- ave(seq_len(nrow(fisher)), fisher$id, FUN = seq_along)

cat("Subjects:", length(unique(fisher$id)),
    "| Total observations:", nrow(fisher), "\n")
cat("Median beeps per subject:",
    median(as.integer(table(fisher$id))), "\n")
for (y in y_fisher) {
  cat(sprintf(" %-25s M = %5.1f  SD = %5.1f\n",
              y, mean(fisher[[y]]), sd(fisher[[y]])))
}
```

#### Fit the model

```{r fit-fisher, eval = has_cmdstan, message = FALSE, warning = FALSE}
fit_gauss_f <- bvar(
  id_col      = "id",
  time_col    = "t",
  y_cols      = y_fisher,
  x_cols      = NULL,
  re_cols     = c("Intercept"),
  re_temporal = FALSE,
  K           = 1,
  data        = fisher,
  family      = "gaussian",
  iter        = 4000,
  warmup      = 1000,
  chains      = 4,
  cores       = 4,
  seed        = 42,
  save_data   = TRUE
)
```

```{r summary-fisher, eval = has_cmdstan}
summary(fit_gauss_f)
```

#### Observed vs predicted

```{r fisher-scatter, eval = has_cmdstan, fig.height = 4}
pred_f <- predict(fit_gauss_f, type = "response")
obs_f  <- fit_gauss_f$data_used
cmp_f  <- merge(obs_f, pred_f, by = c("id", "t"))

par(mfrow = c(1, 3), mar = c(4, 4, 2.5, 1))
for (y in y_fisher) {
  observed  <- cmp_f[[y]]
  predicted <- cmp_f[[paste0("predicted_", y)]]
  r2        <- cor(observed, predicted, use = "complete.obs")^2
  plot(observed, predicted,
       xlab = paste("Observed", y), ylab = paste("Predicted", y),
       main = paste0(y, "\nR\u00b2 = ", round(r2, 2)),
       pch = 16, cex = 0.4, col = rgb(0.2, 0.4, 0.8, 0.3),
       cex.main = 0.9)
  abline(0, 1, col = "firebrick", lwd = 2)
}
```

#### Residuals

```{r fisher-residuals, eval = has_cmdstan, fig.height = 4}
par(mfrow = c(1, 3), mar = c(4, 4, 2.5, 1))
for (y in y_fisher) {
  resid <- cmp_f[[y]] - cmp_f[[paste0("predicted_", y)]]
  hist(resid, breaks = 30, freq = FALSE,
       xlab = "Residual (observed \u2212 predicted)",
       main = paste0(y, " residuals"),
       col = "steelblue3", border = "white",
       cex.main = 0.9)
  curve(dnorm(x, mean = 0, sd = sd(resid, na.rm = TRUE)),
        col = "firebrick", lwd = 2, add = TRUE)
}
```

#### Prediction intervals

We construct 95% prediction intervals as $\widehat y \pm 1.96 \sqrt{\mathrm{SD}_{\mu}^{2}
+ \hat\sigma_{\varepsilon}^{2}}$, combining posterior uncertainty in the
conditional mean (`_sd` columns from `method = "posterior-sample"`) with the
posterior-mean residual SD (`extract_param(fit)`).

```{r fisher-pi, eval = has_cmdstan, fig.height = 4}
pred_f_s <- predict(fit_gauss_f, type = "response",
                    method = "posterior-sample", ndraws = 200, seed = 42)

ptab    <- extract_param(fit_gauss_f)
sigma_f <- setNames(ptab$mean[ptab$type == "sigma"],
                    ptab$outcome[ptab$type == "sigma"])

subj_pi <- names(which.max(table(pred_f_s$id)))
cmp_pi  <- merge(
  obs_f[obs_f$id == subj_pi, c("id", "t", y_fisher)],
  pred_f_s[pred_f_s$id == subj_pi, ],
  by = c("id", "t")
)
cmp_pi <- cmp_pi[order(cmp_pi$t), ]

par(mfrow = c(1, 3), mar = c(4, 4, 2.5, 1))
for (y in y_fisher) {
  mu_hat <- cmp_pi[[paste0("predicted_", y)]]
  sd_mu  <- cmp_pi[[paste0("predicted_", y, "_sd")]]
  pi_w   <- 1.96 * sqrt(sd_mu^2 + sigma_f[[y]]^2)
  lo     <- mu_hat - pi_w
  hi     <- mu_hat + pi_w

  plot(NULL, xlim = range(cmp_pi$t), ylim = range(c(lo, hi, cmp_pi[[y]]), na.rm = TRUE),
       xlab = "Beep (within subject)", ylab = y,
       main = paste0(y, "\nparticipant ", subj_pi),
       cex.main = 0.85)
  polygon(c(cmp_pi$t, rev(cmp_pi$t)), c(lo, rev(hi)),
          col = rgb(0.2, 0.4, 0.8, 0.20), border = NA)
  lines(cmp_pi$t, mu_hat, col = "steelblue4", lwd = 2)
  points(cmp_pi$t, cmp_pi[[y]], pch = 16, cex = 0.6, col = "black")

  cov95 <- mean(cmp_pi[[y]] >= lo & cmp_pi[[y]] <= hi, na.rm = TRUE)
  mtext(sprintf("95%% PI coverage: %.1f%%", 100 * cov95),
        side = 3, line = 0.1, cex = 0.7, col = "grey30")
}
```

The in-sample R² and residual diagnostics are optimistic upper bounds;
out-of-sample evaluation would give more conservative estimates of predictive
performance.

---

## References

Fisher, A. J., Bosley, H. G., Fernandez, K. C., Reeves, J. W., Soyster, P. D.,
Diamond, A. E., & Barkin, J. (2019). Open trial of a personalized
modular treatment for mood and anxiety. *Behaviour Research and Therapy*,
116, 69–79. <https://doi.org/10.1016/j.brat.2019.01.010>

Geschwind, N., Peeters, F., Drukker, M., van Os, J., & Wichers, M. (2011).
Mindfulness training increases momentary positive emotions and reward
experience in adults vulnerable to depression: A randomized controlled trial.
*Journal of Consulting and Clinical Psychology*, 79(5), 618–628.
<https://doi.org/10.1037/a0024595>

Siepe, B. S. (2025). openESM: Access the Open Experience Sampling Method
Database. R package. <https://openesmdata.org>
