---
title: "Prediction"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Prediction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


```{r setup}
library(bvarnet)

# subject to be removed again...
library(qgraph)
```

## Load Data

```{r}
data(studentlife)
```

## Fit the model

To show the full capabilities of the prediction module, we will use the multilevel VAR model as example:

```{r, eval = has_cmdstan}
fit <- bvar(
  id_col = "id",
  time_col = "day",
  y_cols = c("anxious", "calm", "conventional", "critical", "dependable"),
  x_cols = NULL,
  re_cols = c("Intercept"),
  re_temporal = TRUE,
  K = 1,
  data = studentlife,
  family = c("ordinal"),
  iter = 4000,
  warmup = 1000,
  chains = 4,
  cores = 4,
  seed = 1337,
  save_data = TRUE
)
```

```{r}
summary(fit)
```

## 2 — In-Sample Fitted Values

The simplest call returns one-step-ahead fitted values on the linear predictor scale for the training data:

```{r}
p <- predict(fit)
p
```




This returns a **data.frame** with one row per modeled observation (i.e. only the rows that Stan actually used for estimation). The columns include the subject ID, time point, and one `predicted_<y>` column per outcome node.

## 3 — Calibration: Predicted Probabilities by Observed Category

For ordinal outcomes a scatter plot of predicted vs observed values is not meaningful, because the observed values are discrete categories. Instead, we can check whether the model is well-calibrated by asking: *when a response of category k was actually observed, did the model assign the most probability mass to category k?*

We use `type = "probabilities"` which returns a named list of data.frames — one per outcome variable — with columns `cat_1` through `cat_5` containing the predicted probability for each ordinal category.

```{r, fig.width = 7, fig.height = 8}
probs <- predict(fit, type = "probabilities")
obs   <- fit$data_used

y_cols   <- c("anxious", "calm", "conventional", "critical", "dependable")
n_cats   <- 5
cat_cols <- paste0("cat_", seq_len(n_cats))

par(mfrow = c(3, 2), mar = c(4, 4, 2, 1))
for (y in y_cols) {
  cmp <- merge(obs[, c("id", "day", y)], probs[[y]], by = c("id", "day"))

  # For each observed category, compute mean predicted probability per category
  cal <- matrix(NA, n_cats, n_cats,
                dimnames = list(observed = seq_len(n_cats),
                                predicted = cat_cols))
  for (k in seq_len(n_cats)) {
    rows_k <- cmp[[y]] == k
    if (any(rows_k)) {
      cal[k, ] <- colMeans(cmp[rows_k, cat_cols, drop = FALSE])
    }
  }

  barplot(
    t(cal),
    beside = FALSE,
    col    = hcl.colors(n_cats, "Blues 3"),
    xlab   = "Observed category",
    ylab   = "Mean predicted probability",
    main   = y,
    ylim   = c(0, 1)
  )
  legend("topright", legend = rev(seq_len(n_cats)), title = "Predicted",
         fill = rev(hcl.colors(n_cats, "Blues 3")), cex = 0.7, bty = "n")
}
```

In a well-calibrated model, the darkest bar segment (matching the observed category) should dominate each column. That is, when we observe category 3, the model should assign the highest predicted probability to category 3.

## 4 — Individual Probability Trajectory

A particularly informative view for intensive longitudinal data is to look at how the model's predicted probabilities evolve over time for a single participant. This shows both the predicted category and the model's uncertainty at each time point.

```{r, fig.width = 7, fig.height = 5}
probs <- predict(fit, type = "probabilities")
obs   <- fit$data_used

# Pick one subject and one variable
subj <- obs$id[1]
y    <- "anxious"

cmp <- merge(
  obs[obs$id == subj, c("id", "day", y)],
  probs[[y]][probs[[y]]$id == subj, ],
  by = c("id", "day")
)
cmp <- cmp[order(cmp$day), ]

n_cats   <- 5
cat_cols <- paste0("cat_", seq_len(n_cats))
cols     <- hcl.colors(n_cats, "Blues 3")

# Stacked area: cumulative probabilities
plot(NULL, xlim = range(cmp$day), ylim = c(0, 1),
     xlab = "Day", ylab = "Predicted probability",
     main = paste0(y, " — subject ", subj))

cum <- rep(0, nrow(cmp))
for (k in seq_len(n_cats)) {
  new_cum <- cum + cmp[[cat_cols[k]]]
  polygon(
    c(cmp$day, rev(cmp$day)),
    c(new_cum, rev(cum)),
    col = cols[k], border = NA
  )
  cum <- new_cum
}

# Overlay observed category as points (rescaled to 0–1)
obs_scaled <- (cmp[[y]] - 1) / (n_cats - 1)
points(cmp$day, obs_scaled, pch = 16, cex = 1.2)

legend("topright", legend = rev(seq_len(n_cats)), title = "Category",
       fill = rev(cols), cex = 0.7, bty = "n")
```

The shaded areas show the predicted probability for each category at every time point, and the black dots mark the actually observed category (rescaled to the 0–1 range). Time points where a dot sits inside the band of its corresponding category indicate good predictions.

## 5 — Observed vs Predicted Frequencies Over Time

Inspired by Blanken et al. (2019, JAMA Psychiatry), we can compare the observed and predicted category frequencies at each time point, aggregated across all subjects. For each day, the observed frequency is the proportion of subjects in each category, while the predicted frequency is the mean predicted probability across subjects. If the model captures the temporal dynamics well, the two should track closely.

```{r, fig.width = 7, fig.height = 8}
probs <- predict(fit, type = "probabilities")
obs   <- fit$data_used

y_cols   <- c("anxious", "calm", "conventional", "critical", "dependable")
n_cats   <- 5
cat_cols <- paste0("cat_", seq_len(n_cats))
cols     <- hcl.colors(n_cats, "Blues 3")

par(mfrow = c(3, 2), mar = c(4, 4, 2, 1))
for (y in y_cols) {
  cmp <- merge(obs[, c("id", "day", y)], probs[[y]], by = c("id", "day"))
  days <- sort(unique(cmp$day))

  # Compute observed proportions and predicted means per day
  obs_freq  <- matrix(NA, length(days), n_cats)
  pred_freq <- matrix(NA, length(days), n_cats)

  for (i in seq_along(days)) {
    rows_d <- cmp$day == days[i]
    observed_cats <- cmp[[y]][rows_d]
    for (k in seq_len(n_cats)) {
      obs_freq[i, k]  <- mean(observed_cats == k)
      pred_freq[i, k] <- mean(cmp[[cat_cols[k]]][rows_d])
    }
  }

  plot(NULL, xlim = range(days), ylim = c(0, max(obs_freq, pred_freq, na.rm = TRUE)),
       xlab = "Day", ylab = "Frequency", main = y)

  for (k in seq_len(n_cats)) {
    lines(days, obs_freq[, k],  col = cols[k], lwd = 2, lty = 1)
    lines(days, pred_freq[, k], col = cols[k], lwd = 2, lty = 2)
  }

  legend("topright", legend = c(seq_len(n_cats), "Obs.", "Pred."),
         col = c(cols, "black", "black"), lty = c(rep(NA, n_cats), 1, 2),
         pch = c(rep(15, n_cats), NA, NA), lwd = 2, cex = 0.6, bty = "n",
         ncol = 2)
}
```

Solid lines show the observed category frequencies and dashed lines show the model-predicted frequencies at each day. Each color represents one ordinal category.

## 6 — Comparing Predicted vs Observed (Gaussian Example)

For continuous outcomes, we can directly compare predicted fitted values against observed data. We fit a small Gaussian VAR with four variables and use `save_data = TRUE` so the cleaned estimation data is stored in `fit_gauss$data_used`:

```{r, eval = has_cmdstan}
fit_gauss <- bvar(
  id_col = "id",
  time_col = "day",
  y_cols = c("act_running_ep_0", "act_still_ep_0", "act_walking_ep_0"),
  x_cols = NULL,
  re_cols = c("Intercept"),
  re_temporal = FALSE,
  K = 1,
  data = studentlife,
  family = "gaussian",
  iter = 2000,
  warmup = 1000,
  chains = 4,
  cores = 4,
  seed = 1337,
  save_data = TRUE
)
```

```{r, fig.width = 7, fig.height = 6}
pred <- predict(fit_gauss, type = "response")
obs  <- fit_gauss$data_used

cmp <- merge(obs, pred, by = c("id", "day"))

y_cols <- c("act_running_ep_0", "act_still_ep_0", "act_walking_ep_0")

par(mfrow = c(2, 2), mar = c(4, 4, 2, 1))
for (y in y_cols) {
  observed  <- cmp[[y]]
  predicted <- cmp[[paste0("predicted_", y)]]

  plot(
    observed, predicted,
    xlab = "Observed", ylab = "Predicted",
    main = y,
    pch = 16, cex = 0.4, col = rgb(0, 0, 0, 0.25)
  )
  abline(0, 1, col = "red", lwd = 2)
}
```

