---
title: "Random-Effects"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Random-Effects}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

First, we have to load the package:

```{r setup}
library(bvarnet)

# subject to be removed again...
library(qgraph)
```

First, we have to load the data
```{r}
data(studentlife)
```

There is some missing data in the dataset. The models default options handle this by themselves. For a further elaboration on this, you can read `vignette("Missing-Data")`.

Currently, `bvarnet` allowes for the following outcome variables:

- binary (`family = "bernoulli")`
- ordinal (`family = "ordinal")` (adjacent category model)
- continuous (`family = "gaussian")`

For the following tutorial we will investigate the temporal network for the variables `anxious`, `calm`, `conventional`, `critical`, and `dependable`. All these variables are on a ordinal scale. Therefore we will use the adjacent category ordinal model to estimate this model.

First we need to specify the prior distributions that we want to use on the parameters.
For this we can use the `set_priors()` function. As we are running a ordinal model, without covariates, we have to specify priors on the temporal structure (denoted by the $\Phi$ matrix) and on the category threshold parameters ($\kappa$) which are specific to the adjacent category ordinal model.

```{r}
priors <- set_priors(phi = prior(family = "normal",
                                 loc = 0,
                                 scale = 0.5),
                     kappa = prior(family = "normal",
                                   loc = 0,
                                   scale = 2),
                     sigma = prior(family = "normal",
                                   loc = 0,
                                   scale = 3))

priors
```


```{r, eval = has_cmdstan}
fit <- bvar(
  id_col = "id",
  time_col = "day",
  y_cols = c("anxious", "calm", "conventional", "critical", "dependable"),
  x_cols = NULL,
  re_cols = NULL,
  re_temporal = FALSE,
  K = 1,
  data = studentlife,
  family = c("ordinal"),
  priors = priors,
  iter = 4000,
  warmup = 1000,
  chains = 4,
  cores = 4,
  seed = 1337
)
```
